---
title: "Materials for the course"
author: "Alex Sanchez"
date: "June 2019"
output: 
  html_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA, prompt = TRUE, tidy = FALSE, fig.width = 7, fig.height = 7,echo = TRUE, message = FALSE, warning = FALSE, cache=FALSE)
Sys.setlocale("LC_TIME", "C")
```

# Presentation

This is a course on web scraping **with R**. THis means that it is expected that participants in the course are familiar with R, honestly, the more the better. 

The course intends to relies on this familiarity to introduce distinct aspects of web scraping.

# Materials for the course

- It is assumed that the last version of R is installed in your computer.

- The course is intended to be taken using R studio
    + Download and install Rstudio from its web site: [https://www.rstudio.com/](https://www.rstudio.com/)
    
- The course uses control version with git and github
    + Create a free github account at [github.com](github.com) if you don't have one yet.
    + Install te `git` program in your computer. Make sure that git is in the path so that Rstudio is able to locate it.
    
- We will need a twitter account for a part of the course:
    + Create a twitter account at [twitter.com](twitter.com)
    
# R packages

A variety of R packages will be used during the course. Install them in your system using, if you wish, the code below.

```{r eval=FALSE}
# Rlibraries to be used in the Web Scraping course
if(!require(rvest)) install.packages("rvest")
if(!require(plyr)) install.packages("plyr")
if (!require(twitteR)) install.packages("twitteR")
if (!require(ROAuth)) install.packages("ROAuth")
if (!require(tm)) install.packages("tm")
if (! require(wordcloud) ) install.packages("wordcloud")
if (! require(XML) ) install.packages("XML")
###
if(!require(maps)) install.packages("maps")

```

