---
title: "Web Scraping with R (5): <br>Introduction to JSON"
author: "Alex Sanchez (asanchez@ub.edu) <br> GME Department. Universitat de Barcelona <br> Statistics and Bioinformatics Unit. Vall d'Hebron Institut de Recerca"
date: "April 2019"
output:
  slidy_presentation:
    fig_width: 7
    fig_height: 6
css: myStyles.css
footer: "JSON"
keep_md: true
highlight: pygments
editor_options: 
  chunk_output_type: console
---

```{r setLicense, child = 'license.Rmd'}
```

```{r disclaimer, child = 'disclaimer.Rmd'}
```


What is JSON
=============

- JSON stands for *JavaScript Object Notation*
- It is a **format** for representing data.

    + general purpose format
    + lightweight format
    + widely popular
    + fairly simple

- JSON has become the option of choice to provide data for many applications.

JSON structure
===============

- JSON is formed in a javascript-like style (*what doesn't mean much when one is not familiar with javascript*).
- We distinguish Data Types and Data Containers

- JSON Data Types
    + `null`, `true`, `false`, `number`,`string`
    
- JSON Data Containers
    + square brackets `[ ]`, 
    + curly brackets `{ }`