---
title: "Case study on Automated Data Collection"
author: "SIMON MUNZERT, CHRISTIAN RUBBA, PETER MEISSNER, DOMINIC NYHUIS"
date: "June 2018"
output: 
  html_document:
    toc: true
---

This case study has been adapted from chapter 1 of the book [Automated Data Collection with R](http://www.r-datacollection.com/) (ADCR, from now on)

```{r setup, include=FALSE, echo=FALSE}
require(knitr)
# include this code chunk as-is to set options
opts_chunk$set(comment = NA, prompt = TRUE, tidy = FALSE, fig.width = 7, fig.height = 7,echo = TRUE, message = FALSE, warning = FALSE, cache=FALSE)
Sys.setlocale("LC_TIME", "C")
```

```{r}
# load packages
if (!require(stringr)) install.packages("stringr", dep=TRUE)
if (!require(XML)) install.packages("XML", dep=TRUE)
if (!require(maps)) install.packages("maps", dep=TRUE)

require(stringr)
require(XML)
require(maps)
```

# Getting the data

We start grabbing the data from the web page [https://en.wikipedia.org/wiki/List_of_World_Heritage_in_Danger](https://en.wikipedia.org/wiki/List_of_World_Heritage_in_Danger).

If there are any problems the page can be saved into a local file and scraping performed on a local copy.

```{r}
# parsing from HTML file in the web
# heritage_parsed <- htmlParse("https://en.wikipedia.org/wiki/List_of_World_Heritage_in_Danger", encoding = "UTF-8")

# parsing from locally stored HTML file
heritage_parsed <- htmlParse("worldheritagedanger.htm")
```

## Parsing the object
We have obtained an object with all that's in the web page.
Next we extract *HTML tables* from this object.

```{r}
tables <- readHTMLTable(heritage_parsed, stringsAsFactors = FALSE)
names(tables)
# extract desired table
danger_table <- tables[[2]]

# alternatively: directly select second table
danger_table <- readHTMLTable(heritage_parsed, stringsAsFactors = FALSE, which = 2) 

names(danger_table)
```

# Data cleansing

Now we have a data.frame that can be explored "as usual" 
```{r}
# select and rename columns
danger_table <- danger_table[,c(1,3,4,6,7)]
colnames(danger_table) <- c("name","locn","crit","yins","yend")
danger_table$name[1:3]
```
Additionally, we perform some simple data cleaning, a step
often necessary when importing web-based content into R. 

- The variable crit, which contains the information whether the site is of cultural or natural character, is recoded, 
- and the two variables y_ins and y_end are turned into numeric ones.2 
- Some of the entries in the y_end variable are ambiguous as they contain several years. We select the last given year in the cell. 
- To do so, we specify a so-called regular expression 

```{r}
danger_table$crit <- ifelse(str_detect(danger_table$crit, "Natural")==T, "nat", "cult")
# cleanse years
danger_table$yins <- as.numeric(danger_table$yins)
danger_table$yend
yend_clean <- unlist(str_extract_all(danger_table$yend, "[[:digit:]]{4}$"))
danger_table$yend <- as.numeric(yend_clean)
```
The locn variable is a bit of a mess, exemplified by three cases drawn from the data-set:
```{r}
danger_table$locn[c(1,3,5)]
```
The variable contains the name of the site's location, the country, and the geographic coordinates in several varieties. What we need for the map are the coordinates, given by the latitude (e.g., 30.84167N) and longitude (e.g., 29.66389E) values. 

To extract this information, we have to use some more advanced text manipulation tools called "regular expressions".

In short, we have to give R an exact description of what the information we are interested in looks like, and then let R search for and extract it. 

To do so, we use functions from the stringr package. In order to get the latitude and longitude values, we write the following:

```{r}
# get countries
reg <- "[[:alpha:] ]+(?=[[:digit:]])"
country <- str_extract(danger_table$locn, perl(reg)) 
# use forward assertion in Perl regular expression
country
country[29] <- "C?te d'Ivoire / Guinea"
country[32] <- ""
danger_table$country <- country

# get coordinates
reg_y <- "[/][ -]*[[:digit:]]*[.]*[[:digit:]]*[;]"
reg_x <- "[;][ -]*[[:digit:]]*[.]*[[:digit:]]*"
y_coords <- str_extract(danger_table$locn, reg_y)
(y_coords <- as.numeric(str_sub(y_coords, 3, -2)))
danger_table$y_coords <- y_coords
x_coords <- str_extract(danger_table$locn, reg_x)
(x_coords <- as.numeric(str_sub(x_coords, 3, -1)))
danger_table$x_coords <- x_coords
danger_table$locn <- NULL
```
When we apply this pattern to the locn variable with the
str_extract() command and extract the numeric information with str_sub(), we get the following:

```{r}
round(danger_table$y_coords, 2)[1:3]
round(danger_table$x_coords, 2)[1:3]
dim(danger_table)
head(danger_table)
```
# Data visualization

The data frame consists of 44 observations and 6 variables. The data are now set up in a way that we can proceed with mapping the sites. To do so, we use another package named "maps." In it we find a map of the world that we use to pinpoint the sites' locations with the extracted y and x coordinates.

```{r}
# plot endangered heritage sites
# pdf(file="heritage-map.pdf", height=3.3, width=7, family="URWTimes")
par(oma=c(0,0,0,0))
par(mar=c(0,0,0,0))
pch <- ifelse(danger_table$crit == "nat", 19, 2)
map("world", col = "darkgrey", lwd = .5, mar = c(0.1,0.1,0.1,0.1))
points(danger_table$x_coords, danger_table$y_coords, pch = pch, col = "black", cex = .8)
box()
# dev.off()
```

We find that many of the endangered sites are located in Africa, the Middle East, and Southwest Asia, and a few others in South and Central America. The endangered cultural heritage sites are visualized as the triangle. They tend to be clustered in the Middle East and Southwest Asia. Conversely, the natural heritage sites in danger, here visualized as the dots, are more prominent in Africa.

We find that there are more cultural than natural sites in danger.

```{r}
# table heritage criteria
table(danger_table$crit)
```

Other plots can be done such as year of endangerment
```{r}
#pdf(file="heritage-hist1.pdf", height=3.3, width=7, family="URWTimes")
par(oma=c(0,0,0,0))
par(mar=c(4,4,1,.5))
hist(danger_table$yend, freq=TRUE, xlab="Year when site was put on the list of endangered sites", main="")
box()
#dev.off()
```

We can plot time between inscription and endangerment
 
```{r}
duration <- danger_table$yend - danger_table$yins
#pdf(file="heritage-hist2.pdf", height=3.3, width=7, family="URWTimes")
par(oma=c(0,0,0,0))
par(mar=c(4,4,1,.5))
hist(duration, freq=TRUE, xlab="Years it took to become an endangered site", main="")
box()
#dev.off()
```



