---
title: "Exercises on regular expressions"
output: 
  html_document:
    toc: true
    float: true
---

```{r setup, include=FALSE, echo=FALSE}
require(knitr)
# include this code chunk as-is to set options
opts_chunk$set(comment = NA, prompt = TRUE, tidy = FALSE, fig.width = 7, fig.height = 7, echo = FALSE, message = FALSE, warning =FALSE, cache=FALSE, print=FALSE, results=FALSE, include=FALSE)
Sys.setlocale("LC_TIME", "C")
```

```{r, echo=FALSE, message=FALSE}
# load packages
if (!require(stringr)) install.packages("stringr", dep=TRUE)
require(stringr)
```

## Exercise 1

- Create a variable called `text1` and populate it with the value “The current year is 2019”
- Create a variable called `my_pattern` and implement the required pattern for finding *any digit* in the variable `text1`.
- Use function `grepl` to verify if there is a digit in the string variable.

```{r exercise1}
text1 <- "The current year is 2019 but next one will be 2020"
my_pattern <- "[A-z]*[0-9]+[A-z]*"
str_extract_all(text1, my_pattern)
my_pattern2 <-  "[0-9]+"
str_extract_all(text1, my_pattern2)
```


## Exercise 2

- Use function `gregexpr` to find all the positions in `text1` where there is a digit.
- Place the results in a variable called *string_position*
- Can you obtain the same result using a function from the `stringr` package?
```{r exercise2}
string_position <- gregexpr(my_pattern,text1)
string_position[[1]][1:length(string_position[[1]])]
require(stringr)
str_locate(text1, my_pattern)
```

## Exercise 3

- Create a variable called `my_pattern` and implement the required pattern for finding **one digit** and **one uppercase alphanumeric** character, in variable `text1`. HINT: combine predefined classes in the regex pattern.
- Use function `grepl` or its `stringr` equivalent to verify if the searched pattern exists on the string.
```{r exercise3}
my_pattern <- "[[:upper:][:digit:]]"
grepl(my_pattern,text1)
str_locate_all(text1, my_pattern)
```

## Exercise 4
- Use function `regexpr` to find the position of the first space in `text1`.
- Place the results in a variable called `first_space` and Use function `grepl` or its `stringr` equivalent to verify if the searched pattern exists on the string.
```{r exercise4}
my_pattern <- "[[:blank:]]"
first_space <- regexpr(my_pattern,text1)
first_space[[1]][1]
str_locate(text1, my_pattern)
```


## Exercise 5
- Create a pattern that checks in `text1` if there is a lowercase character, followed by any character and then by a digit.
```{r exercise5}
my_pattern <- "[[:lower:]].[[:digit:]]"
grepl(my_pattern,text1)
str_detect(text1, my_pattern)
```


## Exercise 6
- Find the starting position of the above string. Place the results in a variable called `string_pos2`
```{r exercise6}
string_pos2 <- str_locate(text1, my_pattern)
string_pos2[1]
string_pos2 <- gregexpr(my_pattern,text1)[[1]][1]
string_pos2
```

## Exercise 7
- Find the following pattern: one space followed by two lowercase letters and one more space.
- Use a function that returns the starting point of the found string and place its result in `string_pos3`.
```{r exercise7}
my_pattern <- "\\s[a-z][a-z]\\s"
string_pos3 <- str_locate(text1, my_pattern)
string_pos3
string_pos3 <- gregexpr(my_pattern,text1)[[1]][1]
string_pos3
```

## Exercise 8
- Using the sub function, replace the pattern found on the previous exercice by the string " is not ""
- Place the resulting string in `text2` variable.
```{r exercise8}
text2 <- sub(my_pattern," is not ",text1)
text2
text2 <- str_replace(text1, my_pattern," is not ")
text2
```

## Exercise 9
- Find in `text2` the following pattern: Four digits starting at the end of the string.
- Use a function that returns the starting point of the found string and place its result in `string_pos4`.
```{r exercise9}
my_pattern <- "\\d{4}$"
string_pos4 <- gregexpr(my_pattern,text2)[[1]][1]
string_pos4
string_pos4 <- str_locate(text2, my_pattern)
string_pos4
```

## Exercise 10
- Using the `substr` function, and according to the position ofthe string found in the previous excercise, extract the first
two digits found at the end of `text2`.
```{r exercise10}
substr(text2,start = string_pos4,string_pos4+1)
```

## Exercise 11

- File "LipidsData.csv" contains the values obtained in a metabolomics studies on lipidic concentrations in HIV patients.
- The researchers who provided us with the data for the analysis also need to extract some information from the lipid names and give us this information.
    + The nomenclature is easy: We wish you to extract
        - the number of carbon atoms (the first number, before the two points-) 
        - vs. number of double bonds (the second number -after two points-) and also 
        - the lipid family (last part of the name that is not a number.
    + Example
        - C24Cer 24 carbons; 0 double bonds, family name="Cer"
        - C24: 1Cer (a) 24 carbons; 1 double bonds, family name="Cer"
        - C24: 2Cer 24 carbons; Two double bonds, family name="Cer"
        
- Read the file into R and prepare a script that parses the names and writes another file with the information desired.


